Automatically generated NE tagged corpora for English and Hungarian

### 2012-07-01 paper/jeju_presentation.tex
Incorporated Eszter's changes to the presentation.

\newcommand{\vitem}{\vfill \item}
\newcommand{\nyil}{$\rightarrow$\ }

@@

\date % (optional)
{2012.07.12.}

@@

\begin{itemize}
\item highly domain-specific
\item restricted in size
\end{itemize}

@@

\begin{itemize}
\item merging datasets \nyil problems of combining different tagsets
\item manual annotation \nyil not cost-effective
\end{itemize}

@@

\subsection{Named Entity Tagging}
\begin{frame}{Named Entity Tagging}
  The entity list from DBpedia is used for NE tagging.
  \begin{itemize}
  \vitem Entity identification:
    \begin{itemize}
    \vitem Look up linked pages in the entity list
    \vitem References to the current page, if in the list
    \end{itemize}
  \vitem NE tagging: from the DBpedia - CoNLL category mapping
  \vitem Exceptions:
    \begin{itemize}
    \vitem Not entities: common nouns, date expressions, etc.
    \vitem Unmarked entities: link to already identified entities
    \vitem Special cases: derived words, first word in a sentence, etc.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{The Hungarian Corpus}
  The algorithm is, in theory, language-agnostic. We tested this property by
  applying it to a medium-sized language, Hungarian. \\
  Difference from bigger, well-resourced languages:
  \begin{itemize}
  \vitem E % TODO: ide mit?
  \end{itemize}
\end{frame}

\begin{frame}{Differences from English}
  \begin{itemize}
  \vitem Language processing tools
    \begin{itemize}
    \vitem Inflectional morphology: different set of tools were needed 
    \vitem We developed a flexible framework that enables the plugging of any NLP tool
    % TODO: what kind of linguistic tools are needed: table?
    \end{itemize}
  \vitem Wikipedia size and coverage
    \begin{itemize}
    \vitem Hungarian is medium-size language: ~200k articles vs. ~4M English pages
    \vitem Cross-references may point to English pages
    \vitem Templates are in Hungarian: per-language template configuration
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Differences from English -- cont.}
  \begin{itemize}
  \vitem DBpedia coverage
    \begin{itemize}
    \vitem Internationalized chapters for big languages only
    \vitem "Hungarian" DBpedia list contains English articles % Nem egyértelmű
      \begin{itemize}
      \vitem Hungarian equivalents must be added to the list % Added: lásd előző oldal 2/2
      \vitem Hungarian entities without an English page are missing:
             the NER language model might not be accurate
      \vitem Future work: extend the list with the missing Hungarian entities
      \end{itemize}
    \end{itemize}
  \vitem NE conventions
    \begin{itemize}
    \vitem Szeged NER guideline: similar to CoNLL
    \vitem Common nouns in Hungarian: names of languages, religions, months, etc.
    % TODO: compounding
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{itemize}
\item available under CC-BY-SA 3.0 Unported License
\item freely downloadable from {\tt http://hlt.sztaki.hu/resources/} 
\item also distributed through META-SHARE (from 1 August)
\end{itemize}

@@

@@

@@

@@

@@

@@

\section{Conclusions}
\begin{frame}{Results -- cont.}
  Single hold-out vs. cross-validation:
  \begin{itemize}
  \vitem Single hold-out is biased
  \vitem cross-validation gives more stable results
  \end{itemize}
  \vfill
\end{frame}

>>>>

\newcommand{\vitem}{\vfill \item}
\newcommand{\nyil}{$\rightarrow$\ }
\newcommand{\nagytilde}{$\sim$}

@@

\date % (optional)
{2012.07.12. \\ ACL NEWS2012 workshop}

@@

\begin{itemize}
\vitem highly domain-specific
\vitem restricted in size
\end{itemize}
\vfill

@@

\begin{itemize}
\vitem merging datasets \nyil problems of combining different tagsets
\vitem manual annotation \nyil not cost-effective
\end{itemize}
\vfill

@@

\subsection{Named Entity Tagging}
\begin{frame}{Named Entity Tagging}
  The entity list from DBpedia is used for NE tagging.
  \begin{itemize}
  \vitem Entity identification:
    \begin{itemize}
    \vitem Look up linked pages in the entity list
    \vitem References to the current page, if in the list
    \end{itemize}
  \vitem NE tagging: from the DBpedia--CoNLL category mapping
  \vitem Exceptions:
    \begin{itemize}
    \vitem Not entities: common nouns, date expressions, etc.
    \vitem Unmarked entities: link to already identified entities
    \vitem Special cases: derived words, first word in a sentence, etc.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

%\begin{frame}{The Hungarian Corpus}
%  The algorithm is, in theory, language-agnostic. We tested this property by
%  applying it to a medium-sized language, Hungarian. \\
%  Difference from bigger, well-resourced languages:
%  \begin{itemize}
%  \vitem E % TODO: ide mit?
%  \end{itemize}
%\end{frame}

%TODO: the "Hungarian Corpus" phrase must appear in the frame title 
\begin{frame}{Differences from English}
  \begin{itemize}
  \vitem Language processing tools
    \begin{itemize}
    \vitem Inflectional morphology: different set of tools were needed 
    \vitem We developed a flexible framework that enables the plugging of any NLP tool
    % TODO: what kind of linguistic tools are needed: table?
    \end{itemize}
  \vitem Wikipedia size and coverage
    \begin{itemize}
    \vitem Hungarian is medium-size language: \nagytilde 200k articles vs.~\nagytilde 4M English pages
    \vitem Cross-references may point to English pages
    \vitem Templates are in Hungarian: per-language template configuration
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Differences from English -- cont.}
  \begin{itemize}
  \vitem DBpedia coverage
    \begin{itemize}
    \vitem Internationalized chapters for big languages only
    \vitem "Hungarian" DBpedia list contains English articles % Nem egyértelmű
      \begin{itemize}
      \vitem Hungarian equivalents must be added to the list % Added: lásd előző oldal 2/2
      \vitem Hungarian entities without an English page are missing:
             the NER language model might not be accurate
      \vitem Future work: extend the list with the missing Hungarian entities
      \end{itemize}
    \end{itemize}
  \vitem NE conventions
    \begin{itemize}
    \vitem Szeged NER guideline: similar to CoNLL
    \vitem Common nouns in Hungarian: names of languages, religions, months, etc. \nyil more consistent MISC category
    \vitem NEs can be subject of compounding \nyil the joint common noun can modify the original sense of NE: \\ \textit{WorldCom} (\texttt{ORG}) \nyil \textit{WorldCom-botrány} ['WorldCom scandal'] (\texttt{O})
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{itemize}
\item available under CC-BY-SA 3.0 Unported License
\item freely downloadable from {\tt http://hlt.sztaki.hu/resources/} 
\item also distributed through META-SHARE (from August)
\end{itemize}

@@

\begin{frame}{Silver standard corpora}

Silver standard corpora can be useful for improving NER in several ways:

\begin{itemize}
\item for less resourced languages they can serve as training corpora in lieu of gold standard datasets
\item they can serve as training sets for domains differing from newswire
\item they can be sources of huge entity lists
\item they can be sources of feature extraction
\end{itemize}

\begin{frame}{Evaluation}

A maxent NE tagger was used with the feature set consisting of

\begin{itemize}
\item gazetteer features
\item sentence start and end position
\item Boolean-valued orthographic properties of the word form
\item string-valued properties of the word form
\item morphological information
\end{itemize}

The CoNLL standard method was used for evaluation.

@@

%\smallskip

@@

quite low NE density \nyil sentences without NEs were filtered out from the evaluation datasets

\begin{frame}{Experiments}

\begin{nix}
our experiments were performed with a sample of 3.5 million tokens
\end{nix}

\begin{itemize}
\item \textit{CoNLL} train -- \textit{CoNLL} test
\item \textit{enwiki} train -- \textit{enwiki} test
\item \textit{enwiki} train -- \textit{CoNLL} test
\item \textit{CoNLL} train with \textit{wikilists} (gazetteer lists extracted from WP) -- \textit{CoNLL} test
\item \textit{CoNLL} train with \textit{wikitags} (labeling the CoNLL datasets by the model trained on WP, and giving these labels as extra features) -- \textit{CoNLL} test
\end{itemize}

@@

\begin{itemize}
\item training and testing across different corpora decreases F-measure
\item the NE tagger trained on WP does not achieve as high performance tested against CoNLL test set as one trained on its own train set
\item gazetteer lists extracted from WP and training with extra features given by the model trained on WP improve F-measure
\end{itemize}

@@

\begin{itemize}
\item F-measures for Hungarian are higher 
\begin{itemize}
\item the MISC category is more consistent
\item the Szeged NER corpus is highly accurately tagged 
\end{itemize}
\item the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
\end{itemize}

@@

%\begin{frame}{Results -- analysis}

%\begin{itemize}
%\vitem training and testing across different corpora decreases F-measure
%\vitem the NE tagger trained on WP does not achieve as high performance tested against CoNLL test set as one trained on its own train set
%\vitem gazetteer lists extracted from WP and training with extra features given by the model trained on WP improve F-measure
%\vitem F-measures for Hungarian are higher 
%\begin{itemize}
%\vitem the MISC category is more consistent
%\vitem the Szeged NER corpus is highly accurately tagged 
%\end{itemize}
%\end{itemize}
%  \vfill

%\end{frame}

\section{Conclusions}
\begin{frame}{Conclusions}
  \begin{itemize}
 \vitem we presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from WP
  \vitem we applied a new approach: mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to WP entities
  \vitem the huge amount of WP articles opens the possibility of building large enough corpora for less resourced languages such as Hungarian
  \vitem silver standard corpora can improve NER accuracy in more ways
  \vitem the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
  \end{itemize}
  \vfill
\end{frame}

\begin{frame}

{\LARGE Thank you for your attention!}

\bigskip
\bigskip

emails: {\tt simon.eszter@nytud.mta.hu \\ nemeskey.david@sztaki.mta.hu} \\

the corpora are freely downloadable from {\tt http://hlt.sztaki.hu/resources/}
